---
title: "Fundamental QC to check reliability of proteins found in quantitative MS"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

Genes in set 1-2-3 only, list with their names
```{r}
library(tidyverse)
genes_in_1_df <- genes_in_sets[1:5]  %>% drop_na
genes_in_1_list <- as.list(genes_in_1_df[1])

genes_in_2_df <- genes_in_sets[c(1:4, 6)]  %>% drop_na
genes_in_2_list <- as.list(genes_in_2_df[1])

genes_in_3_df <- genes_in_sets[c(1:4, 7)]  %>% drop_na
genes_in_3_list <- as.list(genes_in_3_df[1])

```
Repeat for TMT as well (better, more conservative) USE THESE
```{r}
### List of genes where all samples in set had a valid value (no NA in all sets) - more conservative
```{r}
set1_cols <- grep("Set1", names(genes_w_TMT), value = TRUE)
gene_levels_in_1<- genes_w_TMT[, c('Gene.Name', 'Protein.ID.s.', set1_cols)]  %>% drop_na   # 5696

set2_cols <- grep("Set2", names(genes_w_TMT), value = TRUE)
gene_levels_in_2<- genes_w_TMT[, c('Gene.Name', 'Protein.ID.s.', set2_cols)]  %>% drop_na  # 1544

set3_cols <- grep("Set3", names(genes_w_TMT), value = TRUE)
gene_levels_in_3<- genes_w_TMT[, c('Gene.Name', 'Protein.ID.s.', set3_cols)]  %>% drop_na   # 1916
```




Find genes shared by all and create a list w their names
```{r}
library(tidyr)
genes_in_123_df <- genes_in_sets %>% drop_na
genes_in_123_list <- as.list(genes_in_123_df[1])
```

List of genes only present in set1

```{r}

all_genes <-  c(genes_in_1_list$Gene.Name, genes_in_2_list$Gene.Name, genes_in_3_list$Gene.Name) %>% unique()
genes_only_set1 <- all_genes[!all_genes %in% genes_in_123_list$Gene.Name]   # it's a list
```


Venn diagram method 1 (ugly but informative)
```{r}
library(ggVennDiagram)
x <- list(A = genes_in_1_list$Gene.Name, B = genes_in_2_list$Gene.Name, C = genes_in_3_list$Gene.Name)
ggVennDiagram(x, set_label= c("Set 1", "Set 2", "Set 3"))
```

Venn using ggplot2
```{r}
library(ggplot2)
venn = Venn(list(genes_in_1_list$Gene.Name, genes_in_2_list$Gene.Name, genes_in_3_list$Gene.Name))
ggvenn(venn, thickness = 0, alpha = 0.3, slice = "all")

```

Save lists into txt file for DAVID
```{r}

write.table(genes_in_3_list$Gene.Name, "genes_in_3.txt", col.names = FALSE, row.names = FALSE, quote = FALSE)
write.table(genes_in_1_list$Gene.Name, "genes_in_1.txt", col.names = FALSE, row.names = FALSE, quote = FALSE)
write.table(genes_in_2_list$Gene.Name, "genes_in_2.txt", col.names = FALSE, row.names = FALSE, quote = FALSE)
write.table(genes_only_set1, "genes_only_in_1.txt", col.names = FALSE, row.names = FALSE, quote = FALSE)
```

# Ari May 2023 - for thesis
```{r}
library(dplyr)
library(ggplot2)
library(janitor)
library(tibble)

```



```{r}

load("C:/Users/1aria/Desktop/rep/Data/Output_data/raw_gene_table.Rda")
genes_w_TMT <- first_run[c(1:4,8:10, 20:118)]
genes_in_sets <- first_run[c(1:4,8:10)]
```


### List of genes with at least 1 peptide count
```{r}
list_genes_in_set_1 <- genes_in_sets %>% column_to_rownames(var = 'Gene.Name') %>% select(Set1_Peptide.count) %>% na.omit() %>% rownames() %>% as.list.data.frame()  %>% unlist()   #5701, as report says! (gene-centric report)

list_genes_in_set_2 <- genes_in_sets %>% column_to_rownames(var = 'Gene.Name') %>% select(Set2_Peptide.count) %>% na.omit() %>% rownames() %>% as.list.data.frame() %>% unlist()    #1551, Report says 1549

list_genes_in_set_3 <- genes_in_sets %>% column_to_rownames(var = 'Gene.Name') %>% select(Set3_Peptide.count) %>% na.omit() %>% rownames() %>% as.list.data.frame()  %>% unlist()  #1923, report says 1921

```

### Venn diagram
```{r}
library(VennDiagram)

venn.diagram(
        x = list(list_genes_in_set_1 , list_genes_in_set_2 , list_genes_in_set_3),
        filename = 'C:/Users/1aria/Desktop/rep/Output/Plots/Venn_sets.png',
        category.names = c("Set 1 (5701)" , "Set 2 (1551) " , "Set 3 (1923)"),
        output = TRUE ,
        disable.logging = TRUE,
        imagetype = 'png',
      lwd = 1,
          col=c("grey77", '#76b8a9', '#993366'),
          fill = c(alpha("grey77",0.7), alpha('#76b8a9',0.7), alpha('#993366',0.7)),
          cex = 1.5,
          fontfamily = "sans",
          cat.cex = 1.1,
          cat.default.pos = "outer",
          cat.fontfamily = "sans",
          cat.col = 'black',
          cat.fontface = 'bold',
          rotation = 1
        )
        

```

### Stacked visual
```{r}
proteins_overlap <- data.frame(Overlap_type = c('In 3 sets', 'In 2 sets', 'In 1 set'), nr_proteins = c(1299, 491, 4296), dumb= c(0, 0, 0))

proteins_overlap$Overlap_type <- factor(proteins_overlap$Overlap_type , levels = c("In 1 set", "In 2 sets", "In 3 sets"))

stacked <- ggplot(proteins_overlap, aes(x = dumb, y = nr_proteins, fill = Overlap_type)) + 
  geom_bar(stat = 'identity', width = 0.5)+
  theme_minimal()+
  xlim(c(-0.3, 1)) +
  labs( title = 'Genecentric proteins', y = 'Gene-centric proteins')+
  theme(
    axis.text.y = element_blank(), 
    axis.text.x = element_text(size = 20), # eliinates x-axis ticks 
    axis.title.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position="bottom")+
    scale_fill_manual(values=c('grey77', '#76b8a9', "#1380A1"))+
  coord_flip()

ggsave(path = "C:/Users/1aria/Desktop/rep/Output/Plots", filename = "Stacked_barchar_protein_overlap.png", stacked, dpi = 300)
  
```
# Confidence in our results: PSM and unique peptides

### Make df with stricter conditions
```{r}
index_PSM_count <- grep("_PSM.count", names(first_run))
index_Unique_peptide_count <- grep("_Unique.peptide.count", names(first_run))
index__Peptide.count <- grep("_Peptide.count", names(first_run))

QC_first_run <- first_run  %>% select(all_of(c(1, index__Peptide.count, index_PSM_count, index_Unique_peptide_count))) %>% column_to_rownames(var= 'Gene.Name')

# save(QC_first_run, file = "C:/Users/1aria/Desktop/rep/Data/Output_data/QC_first_run.Rda")

QC_fist_run_set1 <- QC_first_run %>% select(all_of(grep('Set1', names(QC_first_run))))
QC_fist_run_set1_noNAs <- filter(QC_fist_run_set1, rowSums(is.na(QC_fist_run_set1 )) != ncol(QC_fist_run_set1))  # 5701 IDs

QC_fist_run_set2 <- QC_first_run %>% select(all_of(grep('Set2', names(QC_first_run))))
QC_fist_run_set2_noNAs <- filter(QC_fist_run_set2, rowSums(is.na(QC_fist_run_set2)) != ncol(QC_fist_run_set2))  # 1551 IDs

QC_fist_run_set3 <- QC_first_run %>% select(all_of(grep('Set3', names(QC_first_run))))
QC_fist_run_set3_noNAs <- filter(QC_fist_run_set3, rowSums(is.na(QC_fist_run_set3)) != ncol(QC_fist_run_set3))   # 1923 IDs 

# save(QC_fist_run_set1_noNAs, file = "C:/Users/1aria/Desktop/rep/Data/Output_data/QC_first_run_set1_noNAs.Rda")
# save(QC_fist_run_set2_noNAs, file = "C:/Users/1aria/Desktop/rep/Data/Output_data/QC_first_run_set2_noNAs.Rda")
# save(QC_fist_run_set3_noNAs, file = "C:/Users/1aria/Desktop/rep/Data/Output_data/QC_first_run_set3_noNAs.Rda")

# How many IDs if we restrict to PSM count above 1, Unique peptide above 1, and both? - SET1 (5701)
QC_fist_run_set1_noNAs_strictPSM <- filter(QC_fist_run_set1_noNAs, Set1_PSM.count > 1) # 4417
QC_fist_run_set1_noNAs_strictUnique <- filter(QC_fist_run_set1_noNAs, Set1_Unique.peptide.count > 1) # 3890
QC_fist_run_set1_noNAs_strictUnique_strictPSM <- filter(QC_fist_run_set1_noNAs, Set1_Unique.peptide.count > 1, Set1_PSM.count > 1 ) # 3890

# How many IDs if we restrict to PSM count above 1, Unique peptide above 1, and both? - SET2 (1551)
QC_fist_run_set2_noNAs_strictPSM <- filter(QC_fist_run_set2_noNAs, Set2_PSM.count > 1) # 1192
QC_fist_run_set2_noNAs_strictUnique <- filter(QC_fist_run_set2_noNAs, Set2_Unique.peptide.count > 1) # 1028
QC_fist_run_set2_noNAs_strictUnique_strictPSM <- filter(QC_fist_run_set2_noNAs, Set2_Unique.peptide.count > 1, Set2_PSM.count > 1 ) # 1028


# How many IDs if we restrict to PSM count above 1, Unique peptide above 1, and both? - SET3 (1923)
QC_fist_run_set3_noNAs_strictPSM <- filter(QC_fist_run_set3_noNAs, Set3_PSM.count > 1) # 1460
QC_fist_run_set3_noNAs_strictUnique <- filter(QC_fist_run_set3_noNAs, Set3_Unique.peptide.count > 1) # 1285
QC_fist_run_set3_noNAs_strictUnique_strictPSM <- filter(QC_fist_run_set3_noNAs, Set3_Unique.peptide.count > 1, Set3_PSM.count > 1 ) # 1285

```

### Repeat Venn, on strict PSM dataset
```{r}
strict_list_genes_in_set_1 <- QC_fist_run_set1_noNAs_strictPSM %>% rownames() %>% as.list.data.frame()  %>% unlist() # 4417

strict_list_genes_in_set_2 <- QC_fist_run_set2_noNAs_strictPSM %>% rownames() %>% as.list.data.frame()  %>% unlist()  # 1192 

strict_list_genes_in_set_3 <- QC_fist_run_set3_noNAs_strictPSM %>% rownames() %>% as.list.data.frame()  %>% unlist()  # 1460



venn.diagram(
        x = list(strict_list_genes_in_set_1 , strict_list_genes_in_set_2 , strict_list_genes_in_set_3),
        filename = 'C:/Users/1aria/Desktop/rep/Output/Plots/Venn_PSM_restricted_sets.png',
        category.names = c("Set 1 (4417)" , "Set 2 (1192) " , "Set 3 (1460)"),
        output = TRUE ,
        disable.logging = TRUE,
        imagetype = 'png',
      lwd = 1,
          col=c("grey77", '#76b8a9', '#993366'),
          fill = c(alpha("grey77",0.7), alpha('#76b8a9',0.7), alpha('#993366',0.7)),
          cex = 1.5,
          fontfamily = "sans",
          cat.cex = 1.1,
          cat.default.pos = "outer",
          cat.fontfamily = "sans",
          cat.col = 'black',
          cat.fontface = 'bold',
          rotation = 1
        )
``` 

### Stacked plots
```{r}

Sets <- c(rep("Set 1" , 3) , rep("Set 2" , 3) , rep("Set 3" , 3))
condition <- rep(c("No condition" , "PSM > 1 " , "Unique peptide > 1") , 3)
value <- c(1287, 527, 3890, 359, 163, 1028, 463, 175, 1285)
data <- data.frame(Sets,condition,value)


stacked2 <- ggplot(data, aes(fill=condition, y=value, x=Sets)) + 
    geom_bar(position="stack", stat="identity",  width = 0.5)+
    theme_bw()+
  scale_fill_manual(values=c('grey87', '#76b8a9', "#1380A1"))+
  labs(y = 'Protein IDs', fill = 'Acceptance criteria')+
  theme(
    axis.title.x = element_blank(),
    # axis.text.x = element_text(size = 13),
    # axis.text.y = element_text(size = 13),
    legend.position = 'bottom'
  )

stacked2


ggsave(path = "C:/Users/1aria/Desktop/rep/Output/Plots", filename = "Stacked_barchart_difference_acceptance_criteria.png", stacked2, width=5.5, height = 7, dpi = 300)
```

