---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r}
# Full_sample_info
# info_df
# Metatable

load("C:/Users/1aria/Documents/MTLS/6 - Thesis/04_Analysis/Cleaned_data/longform_info_to_analyze.Rda")
```

```{r}
info_to_analyse_workshop <- t(info_to_analyse) %>% row_to_names(row_number = 1) %>% as.data.frame %>% slice(11:1797) %>% rownames_to_column("Gene")

```

```{r}
longform_info_to_analyze_workshop <- info_to_analyse_workshop %>%
  gather(Sample_name, Count, -Gene) %>% 
  full_join(Metatable, by = "Sample_name")  %>% 
  select(Sample_name, everything()) %>% 
  select(-c(Gene,Count), c(Gene,Count)) 

longform_info_to_analyze_workshop$Count = as.numeric((longform_info_to_analyze_workshop$Count)) 

save(longform_info_to_analyze_workshop, file = "longform_info_to_analyze.Rda")

install.packages("https://cran.r-project.org/src/contrib/Archive/rlang/rlang_0.4.10.tar.gz", repos = NULL, type="source")


install.packages("installr")
library(installr)
updateR()

file.edit(file.path("~", ".Rprofile"))
install.packages("rlang")
install.packages("xfun")

## Boxplot
```{r}

longform_info_to_analyze_workshop %>% filter(Set_nr == 3) %>% 
ggplot() +
  geom_boxplot(mapping = aes(x = Sample_name, y = Count, colour = Sample_nr), show.legend = FALSE)+
  ggtitle('Set 3')+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  scale_color_gradientn(colours = wes_palette("Moonrise2"))
  #+ facet_wrap(~Sample_nr)

```
#Trial at PCA

```{r}
# Packages needed
library(scales)
library(tidyverse)
library(factoextra)

if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install()

BiocManager::install("pcaMethods")
```

```{r}
# Prepare from longform to numerical only matrix
wide_matrix <- longform_info_to_analyze_workshop %>% select(c(Sample_name, Gene, Count)) %>% reshape(idvar = "Sample_name", timevar = "Gene", direction = "wide") %>% as.data.frame() %>% remove_rownames %>%  column_to_rownames(., var = 'Sample_name')

names(wide_matrix) <- gsub("Count.", "", names(wide_matrix))

save(wide_matrix, file = "wide_matrix.RDa")

```
```{r}
pca_t <-  pcaMethods::pca(t(vst), method="bpca", center=TRUE, nPcs = 10)
```

```{r}
pca_PD1 <- prcomp(data=wide_matrix, scale = TRUE, na.action = na.omit)

PD1_BPCA <- pcaMethods::pca(wide_matrix, method="bpca", center=TRUE, nPcs = 10)

plot(pca_t, y = NULL)

```
# 1st attemp Volcano plot
```{r}


```
