---
title: "Normlaity check"
output:
  html_document:
    df_print: paged
  pdf_document: default
editor_options:
  chunk_output_type: inline
fig_width: 15
fig_height: 17
---

Analysis to see if genes are normal (NB: was superfluous analysis, as Nidhi said genes have already been log tranformed previously)

```{r}
library(readxl)
library(dplyr)
library(ggpubr)
library(data.table)
library(tinytex)

set.seed(5)
```

### Preparation: upload the tidied data

```{r}
load("C:/Users/1aria/Documents/MTLS/6 - Thesis/04_Analysis/Cleaned_data/longform_info_to_analyze.Rda")

load("C:/Users/1aria/Documents/MTLS/6 - Thesis/04_Analysis/Cleaned_data/wide_matrix.RDa")

Metatable <- readxl::read_excel("C:/Users/1aria/Documents/MTLS/6 - Thesis/04_Analysis/Raw_data/Metatable.xlsx")
```

### Pre-processing: NAs

I cannot remove complete rows with Nas, or I will not have any data left. Instead, replace NAs with 0

```{r}
wide_matrix_no_NAs <- replace(wide_matrix,is.na(wide_matrix),0) %>% as.data.frame()
```

### 1: what range do my gene levels span?

```{r}
range <- apply(wide_matrix_no_NAs, 2, range) %>% as.data.frame()
range$Minimum <-apply(range,1,FUN=min)
range$Maximum <-apply(range,1, max)
```

*Ans*: From -5.49, to +4.55

### 2: are the gene levels normally distributed?

We analyse that in 3 ways:

-   QQ plots --\> normal if 45° line

-   Density plots --\> should resemble normal disutribution

-   Shapiro Wilks p-value --\> if p-value is below 0.05, then normal

**QQ plot**

```{r}
random_select <- wide_matrix_no_NAs %>% select(sample(seq_len(ncol(.)), size = 9))  %>%  filter(!row_number() %in% c(7,8,9))  # also eliminate ICI rows
```

```{r}
p1 <- ggqqplot(random_select[,1]) + ggtitle(colnames(select(random_select, 1)))
p2 <- ggqqplot(random_select[,2]) + ggtitle(colnames(select(random_select, 2)))
p3 <- ggqqplot(random_select[,3]) + ggtitle(colnames(select(random_select, 3)))
p4 <- ggqqplot(random_select[,4]) + ggtitle(colnames(select(random_select, 4)))
p5 <- ggqqplot(random_select[,5]) + ggtitle(colnames(select(random_select, 5)))
p6 <- ggqqplot(random_select[,6]) + ggtitle(colnames(select(random_select, 6)))
p7 <- ggqqplot(random_select[,7]) + ggtitle(colnames(select(random_select, 7)))
p8 <- ggqqplot(random_select[,8]) + ggtitle(colnames(select(random_select, 8)))
p9 <- ggqqplot(random_select[,9]) + ggtitle(colnames(select(random_select, 9)))

plot <- ggarrange(p1, p2, p3, p4, p5, p6, p7, p8, p9) 

annotate_figure(plot, top = text_grob("QQplots of protein levels across 9 randomly selected genes", 
               color = "#1380A1", face = "bold", size = 14))
```

No, for the most part, the gene counts do not follow a normal distribution. Long horizontal segment signals that many NAs were present in the sample data.

**Density plots**

```{r}
p1 <- ggdensity(random_select[,1]) + ggtitle(colnames(select(random_select, 1)))
p2 <- ggdensity(random_select[,2]) + ggtitle(colnames(select(random_select, 2)))
p3 <- ggdensity(random_select[,3]) + ggtitle(colnames(select(random_select, 3)))
p4 <- ggdensity(random_select[,4]) + ggtitle(colnames(select(random_select, 4)))
p5 <- ggdensity(random_select[,5]) + ggtitle(colnames(select(random_select, 5)))
p6 <- ggdensity(random_select[,6]) + ggtitle(colnames(select(random_select, 6)))
p7 <- ggdensity(random_select[,7]) + ggtitle(colnames(select(random_select, 7)))
p8 <- ggdensity(random_select[,8]) + ggtitle(colnames(select(random_select, 8)))
p9 <- ggdensity(random_select[,9]) + ggtitle(colnames(select(random_select, 9)))

plot <- ggarrange(p1, p2, p3, p4, p5, p6, p7, p8, p9) 

annotate_figure(plot, top = text_grob("Density plots of protein levels across 9 randomly selected genes", 
               color = "#1380A1", face = "bold", size = 14))
```

**Shapiro Wilks test**

```{r}
t <- apply(random_select, 2, shapiro.test)

Shapiro_test <- rbindlist(t, use.names=TRUE, fill=FALSE, idcol=TRUE)

print(Shapiro_test)

```

*Ask Nidhi: is the data ok like this, or should I log transform it?*

### JIC - Log10 transform the data

```{r}
log_transf = function(x) {log10(x+1)}

wide_matrix_logged <- as.data.frame(sapply(wide_matrix_no_NAs, log_transf))

rownames(wide_matrix_logged) <- rownames(wide_matrix_no_NAs)

head(wide_matrix_logged)
```

### JIC - is logged data normal?

```{r}
selected_cols <- colnames(random_select)

random_select_log <- select(wide_matrix_logged, all_of(selected_cols))
```

```{r}
p1 <- ggdensity(random_select_log[,1]) + ggtitle(colnames(select(random_select_log, 1)))
p2 <- ggdensity(random_select_log[,2]) + ggtitle(colnames(select(random_select_log, 2)))
p3 <- ggdensity(random_select_log[,3]) + ggtitle(colnames(select(random_select_log, 3)))
p4 <- ggdensity(random_select_log[,4]) + ggtitle(colnames(select(random_select_log, 4)))
p5 <- ggdensity(random_select_log[,5]) + ggtitle(colnames(select(random_select_log, 5)))
p6 <- ggdensity(random_select_log[,6]) + ggtitle(colnames(select(random_select_log, 6)))
p7 <- ggdensity(random_select_log[,7]) + ggtitle(colnames(select(random_select_log, 7)))
p8 <- ggdensity(random_select_log[,8]) + ggtitle(colnames(select(random_select_log, 8)))
p9 <- ggdensity(random_select_log[,9]) + ggtitle(colnames(select(random_select_log, 9)))

plot <- ggarrange(p1, p2, p3, p4, p5, p6, p7, p8, p9) 

annotate_figure(plot, top = text_grob("Density plots of protein levels across 9 randomly selected genes, after log10 processing", 
               color = "#1380A1", face = "bold", size = 14))
```

**Shapiro Wilks test**

```{r}
t_1 <- apply(random_select_log, 2, shapiro.test)

Shapiro_test_log <- rbindlist(t_1, use.names=TRUE, fill=FALSE, idcol=TRUE)

print(Shapiro_test_log)
```

### Verdict?

Not sure, ask Nidhi.
