---
title: "Differential expression - limma vs t-test"
output: html_notebook
editor_options: 
  chunk_output_type: inline
  markdown: 
    wrap: 72
---

# Find deferentially expressed proteins in our data-set

### The data

-   12 subjects. 6 in each condition (Responders (R), Non-responders
    (NR))
-   All subjects taken at 2x datapoints SCR, C3
-   4x R taken also at C4, C5
-   3x NR taken also at End of treatment
-   35 data points in total

### The comparisons

1.  NR vs R (DONE all)

-   all data points - (Unpaired t-test)

-   SCR only

-   C3 only

-   SCR+C3

2.  Longitudinal (R)

-   SCR vs C3

-   SCR vs C4

-   SCR vs C5

-   C3 vs C4

-   C4 vs C5

3.  Longitudinal (NR)

-   SCR vs C3

-   SCR vs EOT

-   C3 vs EOT

### The methods

-   t-test (paired, unpaired. Basic stat)
-   Limma (normal and robust method)

# The functions

Packages

```{r}
library(dplyr)
library(limma)
library(readxl)
```

Data

```{r}
Metatable_means <- read_excel("C:/Users/1aria/Desktop/rep/Data/Metadata/Metatable_means.xlsx", sheet = "w nur")
load('C:/Users/1aria/Desktop/rep/Data/Output_data/expression_data.Rda')  #expression data, in alphab order. for limma

```

On case you need to make exp again. shouldnt be needing this
```{r}
exp <- data.frame(cbind(names(wide_matrix_t_means), t(wide_matrix_t_means))) %>% select(-c('V1')) %>% tibble::rownames_to_column('Sample_name') %>% filter(Sample_name %in% Metatables_multilevl_comparison$Sample_name) %>% arrange(Sample_name) %>% column_to_rownames(var = 'Sample_name') %>% mutate_if(is.character, as.numeric)

exp_t <- data.frame(cbind(names(exp), t(exp))) %>% select(-c('V1')) %>% mutate_if(is.character, as.numeric)
```


```{r}
# Provide: 
# dataset: numerical only df with genes on rows and samples as col
# colnames_conditionA/B: list containing names of all cols of interest (eg NR vs R)

diff_exp = function(dataset, colnames_conditionA, colnames_conditionB) {
  results = apply(dataset, 1, function(x) t.test(x[colnames_conditionA], x[colnames_conditionB], na.action=na.omit))
  
  colindex_conditionA <-which(colnames(dataset) %in% colnames_conditionA)
  colindex_conditionB  <-which(colnames(dataset) %in% colnames_conditionB)
  
  
  p.val = list()
  for(i in names(results)) {p.val[[i]] = results[[i]]$p.value}
  p_value = as.vector(unlist(p.val))
  
  log2_fc = rowMeans(dataset[colindex_conditionA], na.rm=TRUE)-rowMeans(dataset[colindex_conditionB], na.rm=TRUE)
  
  t_test_result = as.data.frame(cbind(names(results), log2_fc, p_value))
  colnames(t_test_result) = c("Protein", "log2_fc", "p_value")
  
  t_test_result$adj_p_value = p.adjust(t_test_result$p_value, method="fdr")
  
  # t_test_result$Gene_Name = data$X[match(res$Protein, data$geneid)]
  
  t_test_result
}

# Paired t-test
diff_exp_paired = function(dataset, colnames_conditionA, colnames_conditionB) {
  results = apply(dataset, 1, function(x) t.test(x[colnames_conditionA], x[colnames_conditionB], paired = TRUE, na.action=na.omit))
  
  colindex_conditionA <-which(colnames(dataset) %in% colnames_conditionA)
  colindex_conditionB  <-which(colnames(dataset) %in% colnames_conditionB)
  
  
  p.val = list()
  for(i in names(results)) {p.val[[i]] = results[[i]]$p.value}
  p_value = as.vector(unlist(p.val))
  
  log2_fc = rowMeans(dataset[colindex_conditionA], na.rm=TRUE)-rowMeans(dataset[colindex_conditionB], na.rm=TRUE)
  
  t_test_result = as.data.frame(cbind(names(results), log2_fc, p_value))
  colnames(t_test_result) = c("Protein", "log2_fc", "p_value")
  
  t_test_result$adj_p_value = p.adjust(t_test_result$p_value, method="fdr")   # fdr = Benj Hochberg
  
  t_test_result
}
```

# 1. NR vs R

## T-test method first

19 significant proteins (no FC restriction). 16 upregulated (more in NR
than R), 3 downregulated (less in NR than R)

### All datapoints first

```{r}
# Define conditions -->  colnames of R vs NR
NR_colnames <- Metatable_means %>% filter(Sample_status == 'NR') %>% select(Sample_name) %>% as.list() %>% unlist()
R_colnames <- Metatable_means %>% filter(Sample_status == 'R') %>% select(Sample_name) %>% as.list() %>% unlist()

# t-test, unpaired
NR_vs_R_results = as.data.frame(diff_exp(exp_t, colnames_conditionA =NR_colnames, colnames_conditionB =R_colnames ))


##### More stuff #####
# How many significant (no FC constraint)
NR_vs_R_results_sign <- NR_vs_R_results %>% filter(adj_p_value < 0.05)   #19 (16 UP and 3 DOWN)

# UP vs DOWN annotation
NR_vs_R_results_ann <- NR_vs_R_results   %>% 
  na.omit() %>%
  mutate_at(c('log2_fc', 'p_value', 'adj_p_value'), as.numeric) %>%
  mutate(diff_expressed = "NO") %>%
  mutate(diff_expressed = if_else(condition = (log2_fc > 0 & adj_p_value < 0.05) , true = "UP", false = diff_expressed )) %>%
  mutate(diff_expressed = if_else(condition = (log2_fc < 0 & adj_p_value < 0.05) , true = "DOWN", false = diff_expressed ))   %>%
  mutate(label = NA ) %>%
  mutate(label = if_else(condition = (diff_expressed != "NO"), true = Protein, false = label))

NR_vs_R_select <- NR_vs_R_results_ann %>% filter(diff_expressed %in% c('UP','DOWN'))

# save(NR_vs_R_select, file = 'C:/Users/1aria/Desktop/rep/Data/Output_data/Significantly_diff_expressed/ttest_NR_vs_R_significant.Rda')

```

### Only baseline (SCR) points

```{r}
# Define conditions -->  colnames of R vs NR
NR_SCR_colnames <- Metatable_means %>% filter(Sample_status == 'NR', Cycle == 0) %>% select(Sample_name) %>% as.list.data.frame() %>% unlist()
R_SCR_colnames <- Metatable_means %>% filter(Sample_status == 'R', Cycle == 0) %>% select(Sample_name) c %>% unlist()

# t-test, unpaired
NR_vs_R_SCR_results = as.data.frame(diff_exp(exp_t, colnames_conditionA =NR_SCR_colnames, colnames_conditionB =R_SCR_colnames ))

# How many significant (no FC constraint)
NR_vs_R_SCR_results_sign <- NR_vs_R_SCR_results %>% filter(adj_p_value < 0.05)   #0

```

### Only 


### Only C3 points

```{r}
# Define conditions -->  colnames of R vs NR
NR_C3_colnames <- Metatable_means %>% filter(Sample_status == 'NR', Cycle == 1) %>% select(Sample_name) %>% as.list.data.frame() %>% unlist()
R_C3_colnames <- Metatable_means %>% filter(Sample_status == 'R', Cycle == 1) %>% select(Sample_name) %>% unlist()

# t-test, unpaired
NR_vs_R_C3_results = as.data.frame(diff_exp(exp_t, colnames_conditionA =NR_C3_colnames, colnames_conditionB =R_C3_colnames ))

# How many significant (no FC constraint)
NR_vs_R_C3_results_sign <- NR_vs_R_C3_results %>% filter(adj_p_value < 0.05)   #0

```

### SCR + C3

```{r}
# Define conditions -->  colnames of R vs NR
NR_C3_SCR_colnames <- Metatable_means %>% filter(Sample_status == 'NR', Cycle %in% c(0,1)) %>% select(Sample_name) %>% as.list.data.frame() %>% unlist()
R_C3_SCR_colnames <- Metatable_means %>% filter(Sample_status == 'R', Cycle %in% c(0,1)) %>% select(Sample_name) %>% unlist()

# t-test, unpaired
NR_vs_R_C3_SCR_results = as.data.frame(diff_exp(exp_t, colnames_conditionA =NR_C3_SCR_colnames, colnames_conditionB =R_C3_SCR_colnames ))

# How many significant (no FC constraint)
NR_vs_R_C3_SCR_results_sign <- NR_vs_R_C3_SCR_results %>% filter(adj_p_value < 0.05)   #0
```

### Include batch1 in NR (21 samples in NR+batch1, 20 samples in R)
```{r}
# Define conditions -->  colnames of R vs NR
NR_extended_colnames <- Metatable_means %>% filter(Sample_status %in% c('NR', 'Batch1')) %>% select(Sample_name) %>% as.list() %>% unlist()
R_colnames <- Metatable_means %>% filter(Sample_status == 'R') %>% select(Sample_name) %>% as.list() %>% unlist()

# t-test, unpaired
NR_and_batch1_vs_R_results = as.data.frame(diff_exp(exp_t, colnames_conditionA =NR_extended_colnames, colnames_conditionB =R_colnames ))

# How many significant (no FC constraint)
NR_and_batch1_vs_R_results_sign <- NR_and_batch1_vs_R_results %>% filter(adj_p_value < 0.05)   #19

NR_and_batch1_vs_R_results_ann <- NR_and_batch1_vs_R_results   %>% 
  na.omit() %>%
  mutate_at(c('log2_fc', 'p_value', 'adj_p_value'), as.numeric) %>%
  mutate(diff_expressed = "NO") %>%
  mutate(diff_expressed = if_else(condition = (log2_fc > 0 & adj_p_value < 0.05) , true = "UP", false = diff_expressed )) %>%
  mutate(diff_expressed = if_else(condition = (log2_fc < 0 & adj_p_value < 0.05) , true = "DOWN", false = diff_expressed ))   %>%
  mutate(label = NA ) %>%
  mutate(label = if_else(condition = (diff_expressed != "NO"), true = Protein, false = label))

NR_and_batch1_vs_R_select <- NR_and_batch1_vs_R_results_ann  %>% filter(diff_expressed %in% c('UP','DOWN'))

# save(NR_and_batch1_vs_R_select , file = 'C:/Users/1aria/Desktop/rep/Data/Output_data/Significantly_diff_expressed/ttest_NR_and_batch1_vs_R_select .Rda')
```



## Limma

```{r}
Metatable_limma <- Metatable_means %>% select(c('Sample_name', 'Sample_status', 'Cycle', 'Sample_nr')) %>% filter(!Sample_status =='Batch1', !Sample_status =='ICI' ) %>% arrange(Sample_name)

```

### Simplest method: compare 2 groups
```{r}
Groups <- factor(Metatable_limma$Sample_status)
design <- model.matrix(~0+Groups)
colnames(design) <- levels(Groups)

fit<- lmFit(exp_t, design, method = 'robust', maxit=1000)   # method = 'robust', maxit=1000 done by covid nature paper. Omit if you want.

# check this part. How you define this majorly impacts your results
contrast <- makeContrasts( NRvsR = NR-R, levels=design)

fit_contrasts <- contrasts.fit(fit, contrast)
fit_contrasts <- eBayes(fit_contrasts)

print(summary(decideTests(fit_contrasts)))   # 14 DOWN, 37 UP // 23 DOWN and 54 UP if robust method w 1k iterations

topTable(fit_contrasts, coef="NRvsR", number = 67)

#### More stuff
# UP vs DOWN annotation
NR_vs_R_limma_basic_ann <- topTable(fit_contrasts, adjust="fdr", coef="NRvsR", number = 41) %>%
  na.omit %>%
  tibble::rownames_to_column(var = 'Protein') %>%
  mutate(diff_expressed = "NO") %>%
  mutate(diff_expressed = if_else(condition = (logFC > 0) , true = "UP", false = diff_expressed )) %>%
  mutate(diff_expressed = if_else(condition = (logFC < 0) , true = "DOWN", false = diff_expressed )) 

NR_vs_R_limma_robust_ann <- topTable(fit_contrasts, adjust="fdr", coef="NRvsR", number = 57) %>%
  na.omit %>%
  tibble::rownames_to_column(var = 'Protein') %>%
  mutate(diff_expressed = "NO") %>%
  mutate(diff_expressed = if_else(condition = (logFC > 0) , true = "UP", false = diff_expressed )) %>%
  mutate(diff_expressed = if_else(condition = (logFC < 0) , true = "DOWN", false = diff_expressed )) 

# save(NR_vs_R_limma_basic_ann, file = 'C:/Users/1aria/Desktop/rep/Data/Output_data/Significantly_diff_expressed/limma_NR_vs_R_basic_significant' )
# save(NR_vs_R_limma_robust_ann, file = 'C:/Users/1aria/Desktop/rep/Data/Output_data/Significantly_diff_expressed/limma_NR_vs_R_robust_significant' )
```

### SCR only
```{r}
Metatable_limma_SCR <- Metatable_means %>% select(c('Sample_name', 'Sample_status', 'Cycle', 'Sample_nr')) %>% filter(!Sample_status =='Batch1', !Sample_status =='ICI', Cycle == 0 ) %>%  arrange(Sample_name)

exp_t_SCR <- exp_t %>% select(Metatable_limma_SCR$Sample_name)

Groups <- factor(Metatable_limma_SCR$Sample_status)
design <- model.matrix(~0+Groups)
colnames(design) <- levels(Groups)

fit<- lmFit(exp_t_SCR, design, method = 'robust', maxit=1000)   # method = 'robust', maxit=1000 done by covid nature paper. Omit if you want.

# check this part. How you define this majorly impacts your results

contrast <- makeContrasts( NRvsR = NR-R, levels=design)

fit_contrasts <- contrasts.fit(fit, contrast)
fit_contrasts <- eBayes(fit_contrasts)

print(summary(decideTests(fit_contrasts)))   # 0 in basic// 1 UP in robust (SOD2)

### More
# UP vs DOWN annotation

NR_vs_R_SCR_limma_robust_ann <- topTable(fit_contrasts, adjust="fdr", coef="NRvsR", number = 1) %>%
  na.omit %>%
  tibble::rownames_to_column(var = 'Protein') %>%
  mutate(diff_expressed = "NO") %>%
  mutate(diff_expressed = if_else(condition = (logFC > 0) , true = "UP", false = diff_expressed )) %>%
  mutate(diff_expressed = if_else(condition = (logFC < 0) , true = "DOWN", false = diff_expressed ))
         

# save(NR_vs_R_SCR_limma_robust_ann, file = 'C:/Users/1aria/Desktop/rep/Data/Output_data/Significantly_diff_expressed/limma_NR_vs_R_SCR_robust_significant.Rda' )
```

### C3 only
```{r}
# Run whole thing 
Metatable_limma_C3 <- Metatable_means %>% select(c('Sample_name', 'Sample_status', 'Cycle', 'Sample_nr')) %>% filter(!Sample_status =='Batch1', !Sample_status =='ICI', Cycle == 1 ) %>%  arrange(Sample_name)

exp_t_C3 <- exp_t %>% select(Metatable_limma_C3$Sample_name)

Groups <- factor(Metatable_limma_C3$Sample_status)
design <- model.matrix(~0+Groups)
colnames(design) <- levels(Groups)

fit<- lmFit(exp_t_C3, design, method = 'robust', maxit=1000 )   # method = 'robust', maxit=1000 done by covid nature paper. Omit if you want.

# check this part. How you define this majorly impacts your results
contrast <- makeContrasts( NRvsR = NR-R, levels=design)

fit_contrasts <- contrasts.fit(fit, contrast)
fit_contrasts <- eBayes(fit_contrasts)

print(summary(decideTests(fit_contrasts)))   # 0 in basic// 1 UP in robust (SOD2, logFC now more than SCR)

### More
# UP vs DOWN annotation

NR_vs_R_C3_limma_robust_ann <- topTable(fit_contrasts, adjust="fdr", coef="NRvsR", number = 1) %>%
  na.omit %>%
  tibble::rownames_to_column(var = 'Protein') %>%
  mutate(diff_expressed = "NO") %>%
  mutate(diff_expressed = if_else(condition = (logFC > 0) , true = "UP", false = diff_expressed )) %>%
  mutate(diff_expressed = if_else(condition = (logFC < 0) , true = "DOWN", false = diff_expressed ))
         
# save(NR_vs_R_C3_limma_robust_ann, file = 'C:/Users/1aria/Desktop/rep/Data/Output_data/Significantly_diff_expressed/limma_NR_vs_R_C3_robust_significant.Rda' )

```

### SCR + C3
```{r}
# Run whole thing 
Metatable_limma_C3_SCR <- Metatable_means %>% select(c('Sample_name', 'Sample_status', 'Cycle', 'Sample_nr')) %>% filter(!Sample_status =='Batch1', !Sample_status =='ICI', Cycle %in% c(0,1) ) %>%  arrange(Sample_name)

exp_t_C3_SCR <- exp_t %>% select(Metatable_limma_C3_SCR$Sample_name)

Groups <- factor(Metatable_limma_C3_SCR$Sample_status)
design <- model.matrix(~0+Groups)
colnames(design) <- levels(Groups)

fit<- lmFit(exp_t_C3_SCR, design, method = 'robust', maxit=1000 )   # method = 'robust', maxit=1000 done by covid nature paper. Omit if you want.

# check this part. How you define this majorly impacts your results
contrast <- makeContrasts( NRvsR = NR-R, levels=design)

fit_contrasts <- contrasts.fit(fit, contrast)
fit_contrasts <- eBayes(fit_contrasts)

print(summary(decideTests(fit_contrasts)))   # 0 in basic // 17 UP, 9 DOWN ib robust

### More
# UP vs DOWN annotation

NR_vs_R_C3_SCR_limma_robust_ann <- topTable(fit_contrasts, adjust="fdr", coef="NRvsR", number = 26) %>%
  na.omit %>%
  tibble::rownames_to_column(var = 'Protein') %>%
  mutate(diff_expressed = "NO") %>%
  mutate(diff_expressed = if_else(condition = (logFC > 0) , true = "UP", false = diff_expressed )) %>%
  mutate(diff_expressed = if_else(condition = (logFC < 0) , true = "DOWN", false = diff_expressed ))
         
# save(NR_vs_R_C3_SCR_limma_robust_ann, file = 'C:/Users/1aria/Desktop/rep/Data/Output_data/Significantly_diff_expressed/limma_NR_vs_R_C3_robust_significant.Rda' )


```

### For curiosity - NR vs R taking also batch 1, all points
```{r}

# Load extended_exp_t- Expression data, includes batch 1, in order. Generated in 02_process_02_various
load('C:/Users/1aria/Desktop/rep/Data/Output_data/extended_exp_t')

Metatable_limma_batch1 <- Metatable_means %>% select(c('Sample_name', 'Sample_status', 'Cycle', 'Sample_nr')) %>% filter(!Sample_status =='ICI') %>%  arrange(Sample_name)


exp_t_batch1 <- extended_exp_t %>% select(Metatable_limma_batch1$Sample_name)

Groups <- factor(Metatable_limma_batch1$Sample_status)
design <- model.matrix(~0+Groups)
colnames(design) <- levels(Groups)

fit<- lmFit(exp_t_batch1, design, method = 'robust', maxit=1000)   # method = 'robust', maxit=1000 done by covid nature paper. Omit if you want.

# check this part. How you define this majorly impacts your results
contrast <- makeContrasts( NRvsR = (Batch1 + NR) -R, levels=design)

fit_contrasts <- contrasts.fit(fit, contrast)
fit_contrasts <- eBayes(fit_contrasts)

print(summary(decideTests(fit_contrasts)))   # 17 DOWN, 67 UP in basic// 76 DOWN, 125 UP in robust

### More
# UP vs DOWN annotation

NR_and_batch1_vs_R_limma_basic_ann <- topTable(fit_contrasts, adjust="fdr", coef="NRvsR", number = 84) %>%
  na.omit %>%
  tibble::rownames_to_column(var = 'Protein') %>%
  mutate(diff_expressed = "NO") %>%
  mutate(diff_expressed = if_else(condition = (logFC > 0) , true = "UP", false = diff_expressed )) %>%
  mutate(diff_expressed = if_else(condition = (logFC < 0) , true = "DOWN", false = diff_expressed ))


# save(NR_and_batch1_vs_R_limma_basic_ann, file = 'C:/Users/1aria/Desktop/rep/Data/Output_data/Significantly_diff_expressed/limma_NR_and_batch1_vs_R_basic_significant.Rda' )

NR_and_batch1_vs_R_limma_robust_ann <- topTable(fit_contrasts, adjust="fdr", coef="NRvsR", number = 201) %>%
  na.omit %>%
  tibble::rownames_to_column(var = 'Protein') %>%
  mutate(diff_expressed = "NO") %>%
  mutate(diff_expressed = if_else(condition = (logFC > 0) , true = "UP", false = diff_expressed )) %>%
  mutate(diff_expressed = if_else(condition = (logFC < 0) , true = "DOWN", false = diff_expressed ))


save(NR_and_batch1_vs_R_limma_robust_ann, file = 'C:/Users/1aria/Desktop/rep/Data/Output_data/Significantly_diff_expressed/limma_NR_and_batch1_vs_R_robust_significant.Rda' )

```
# 2. Longitudinal (R)
Define paired t-test longitudinal function
```{r}
paired_longitudinal_ttest <- function(Treatment_outcome, chosen_sample_nr, Cycle1, Cycle2) {
  colnames1 <- Metatable_means %>% filter(Sample_status == Treatment_outcome, Cycle == Cycle1) %>% filter(Sample_nr %in% chosen_sample_nr) %>% select(Sample_name) %>% as.list() %>% unlist()
  colnames2 <- Metatable_means %>% filter(Sample_status == Treatment_outcome, Cycle == Cycle2) %>% filter(Sample_nr %in% chosen_sample_nr) %>% select(Sample_name) %>% as.list() %>% unlist()

df = as.data.frame(diff_exp_paired(exp_t, colnames_conditionA =colnames1, colnames_conditionB =colnames2 )) 
}
```


### Comparison
```{r}
# 0: SCR, 1: C3, 2: C4, 3: C5

# SCR vs C3
df_R_SCR_vs_C3 <- paired_longitudinal_ttest('R', 0, 1)   #OP: no significant. Min p-value: 0.89
df_R_SCR_vs_C4 <- paired_longitudinal_ttest('R', chosen_sample_nr = c(1,3,4,6), 0, 2)   #OP: no significant. Min p-value: 0.97
df_R_SCR_vs_C5 <- paired_longitudinal_ttest('R', chosen_sample_nr = c(1,3,4,6), 0, 3)   #OP: no significant. Min p-value: 0.88
df_R_C3_vs_C4 <- paired_longitudinal_ttest('R', chosen_sample_nr = c(1,3,4,6), 1, 2)   #OP: no significant. Min p-value: 0.20
df_R_C3_vs_C5 <- paired_longitudinal_ttest('R', chosen_sample_nr = c(1,3,4,6),  1, 3)   #OP: no significant. Min p-value: 0.32
```

# 3- Longitudinal (NR)
### Comparisons
```{r}
# 0: SCR, 1: C3, 9: EOT

# SCR vs C3
df_NR_SCR_vs_C3 <- paired_longitudinal_ttest('NR', chosen_sample_nr = c(7,8,9, 10, 11, 12), 0, 1)   #OP: no significant. Min p-value: 0.29
df_NR_SCR_vs_EOT <- paired_longitudinal_ttest('NR', chosen_sample_nr = c(7,8,10), 0, 9)  # Does not allow me. 3 points: too few
df_NR_C3_vs_EOT <- paired_longitudinal_ttest('R', chosen_sample_nr = c(7,8,10), 1, 9) # Does not allow me. 3 points: too few.
```
### Save all df for future reference
```{r}
save(df_R_SCR_vs_C3 , file = 'C:/Users/1aria/Desktop/rep/Data/Output_data/Significantly_diff_expressed/ttest_R_SCR_vs_C3 .Rda')
save(df_R_SCR_vs_C4 , file = 'C:/Users/1aria/Desktop/rep/Data/Output_data/Significantly_diff_expressed/ttest_R_SCR_vs_C4 .Rda')
save(df_R_SCR_vs_C5 , file = 'C:/Users/1aria/Desktop/rep/Data/Output_data/Significantly_diff_expressed/ttest_R_SCR_vs_C5 .Rda')
save(df_R_C3_vs_C4 , file = 'C:/Users/1aria/Desktop/rep/Data/Output_data/Significantly_diff_expressed/ttest_R_C3_vs_C4 .Rda')
save(df_R_C3_vs_C5 , file = 'C:/Users/1aria/Desktop/rep/Data/Output_data/Significantly_diff_expressed/ttest_R_C3_vs_C5 .Rda')

save(df_NR_SCR_vs_C3 , file = 'C:/Users/1aria/Desktop/rep/Data/Output_data/Significantly_diff_expressed/ttest_NR_SCR_vs_C3 .Rda')

```

